Community-Conditioned Language Models

### 2020-11-23 paper/paper.tex
various minor revisions

Our dataset includes messages from \num{46} subreddits, 
randomly selected from the set of all forums 
with at least \num{15000} messages per month for each month
in the three years between 2015 and 2017. 
We initially selected \num{50} subreddits, 
but excluded four of them from further analysis: 
two because they were primarily non-English and two with particularly short average message lengths.\footnote{
The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
Each community corpus consist of \num{50000} randomly selected messages from the year 2015.

@@

We experiment with two kinds of model architecture: a simple unidirectional LSTM
\citep{Hochreiter1997} and a Transformer \citep{Vaswani2017}.  In either
case, the model is organised as a standard $n$-layer neural sequence
encoder (we set $n=3$ for all models). As is standard, word tokens are
first converted to vectors using a trainable embedding layer. At the
other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
pre-determined vocabulary.  These encoders form the core of the
community-conditioned language models (CCLM), and we call them
unconditioned models.

To condition the model, we add a \emph{community embedding} parameter to the models.
This parameter is concatenated with the hidden layer of the sequence encoder,
for some layer $l_c \leq n$, and passed through a linear layer 
which projects the resulting vector back to the original hidden layer size.
For $l_c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $l_c=0$, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consists of the most frequent tokens at the message level (across all communities),
meaning that we only count each token once per message.
The main reason for this choice is to avoid over-counting otherwise rare tokens that were spammed
many times in a single message.

For the LSTM, we trained the models on a simple left-to-right language modeling task
with cross-entropy loss over the vocabulary. 
Because the transformer operates on all tokens in the sequence at once,
the language modeling was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion (so we have an auto-regressive\jp{check} model).
We used the AdamW \citep{Loshchilov2019} optimization algorithm,
with an initial learning rate of \num{0.001}, without any extra control on the decay of learning rate.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

We measure the perplexity of each example in the test set. 
For this, we use the negative exponential of the mean cross-entropy loss 
of the language model for a given message and average over a set of messages, $M$.
\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\] \bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?} 
As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models (\cref{tab:model-results}),
but this effect is different in magnitude
for models with different community embedding depths.
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two subsequent LSTM layers. 
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers.

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i \mid m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m \mid c_i)$ (the probability of a message $m$
in the context of community $c_i$) is estimated by the CCLM. 
$P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \sum_i P(c=c_i) P(m\mid c=c_i ). \]
Since the test set is balanced across communities,
we have an even prior for $P(c=c_i)$, giving us
\[P(c=c_i\mid m) \propto  P(m\mid c_i).\]

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

\begin{figure*}
\begin{tikzpicture}
  \confusiongroupplot{lstm-3-1}{transformer-3-3}
\end{tikzpicture}
\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:confusion}
\end{figure*}

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
noted $\Ind_j$. 
First, note that each row of $C$ is a probability distribution,
since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score.
To compute $\Ind_j$, we scale this perplexity by the number of communities, 
which makes $\Ind_j$ comparable across models trained on different
numbers of communities.\footnote{Although we only consider one set of \num{46} communities in this paper.}
\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
Intuitively, $\Ind_j$ is a measure of the uncertainty of the LMCC,
given messages from a community $c_j$.
It ranges from $0$ (for a completely deterministic $C_j$) 
to $1$ (if $C_j$ gives equal weight to every community).

@@

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

\begin{table}
  \centering
  \input{floats/ind_corrs.tex}
  \caption{
    Pearson's $r$ correlation coefficient between community indiscernibility ($\Ind_j$) and
    two different predictors: CCLM perplexity on messages from community $c_j$ ($\Ppl_{M_j}$)
    , and CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).
  }
  \label{tab:ind-corrs}
\end{table}

@@

There are, however, related applications of conditional neural language models.
\citet{Lau2017a} presents a neural language model that jointly learns to predict
words on the sentence-level and represent topics on the document level.
The topic representation is then fed back into the language model, 
improving its performance on next word prediction.
This is similar to how our model improves experiences improved performance
by learning community representations. 
Unlike our model, topics are inferred in an unsupervised way, 
raising the question of whether communities could be identified from 
unlabelled data as well.

Along these lines, \citep{OConnor2010} uses a Bayesian generative model to
infer communities from variation in text data.
In contrast to our work, this model treats words as
independent events, ignoring the structure (and variation) in the
construction of sequences.
It does further suggest, however, that community-level variation
could be modeled in an unsupervised way.

@@

We cannot exclude a correlation (\cref{tab:ind-corrs}) between the perplexity of the
% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46)\jp{make consistent with data once the plot is stablilized}, so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.

@@

The pattern of information gain by community is similar across architectures.
Communities that benefit most from the conditioned model behave that way for 
both the LSTM and transformer.
However, there are some differences.
For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
the best CCLM LSTM but only \num{0.998} for the best transformer model.
Why was the LSTM able to take advantage of the community information for
this community, where the transformer wasn't?
It would be interesting to investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

>>>>

Our dataset includes messages from \num{46} subreddits, 
randomly selected from the set of all forums 
with at least \num{15000} messages per month for each month
in the three years between 2015 and 2017. 
We initially selected \num{50} subreddits, 
but excluded four of them from further analysis: 
two because they were primarily non-English and two with particularly short average message lengths.\\footnote{
The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
% According to your footnote the removed communities are barely outliers. I wouldn't have removed them, but it is too late to do anything about that.
Each community corpus consist of \num{50000} randomly selected messages from the year 2015.

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}.  In either case, the model is organised as a
standard $n$-layer neural sequence encoder (we set $n=3$ for all
models). As conventional, word tokens are first converted to vectors
using a trainable embedding layer. At the other end, word tokens are
predicted using a softmax projection layer. The set of word tokens is
a pre-determined vocabulary.  These encoders form the core of the
of our models, as such do not take community into account, and we call them
unconditioned models.

To condition the model (and obtain community-conditioned language
models --- CCLM), we add a \emph{community embedding} parameter to the
models.  This parameter is concatenated with the hidden layer of the
sequence encoder, for some layer $l_c \leq n$, and passed through a
linear layer which projects the resulting vector back to the original
hidden layer size.  For $l_c = n$, the output of this linear layer is
passed directly to the softmax function, just as the final hidden
layer of the sequence encoder is in other models.  For $l_c=0$, the
community embedding is concatenated with the token embedding.  For
this reason, we set the hidden size of the sequence encoder and the
size of the token to be embedding equal for all models.

@@

We use a vocabulary size of \num{40000} tokens, including a special
out-of-vocabulary token.  The vocabulary consists of the most frequent
tokens at the message level (across all communities), meaning that we
only count each token once per message.  The main reason for this
choice is to avoid over-counting otherwise rare tokens that were
repeated many times in a single message.

For the LSTM, we trained the models on a simple left-to-right language
modeling task with cross-entropy loss over the vocabulary.  Because
the transformer operates on all tokens in the sequence at once, the
language modeling was achieved by masking and incrementally un-masking
input tokens in a left-to-right fashion (so we have an
auto-regressive\jp{check} model).  We used the AdamW
\citep{Loshchilov2019} optimization algorithm, with an initial
learning rate of \num{0.001}, without any extra control on the decay
of learning rate.
%
We used a batch size of \num{256} and a maximum sequence length of
\num{64} tokens, truncating longer messages (16.8\% of messages were
longer than \num{64} tokens).  During training, a dropout rate of
$0.1$ was applied between encoder layers and after each linear layer.

@@

We measure the perplexity of each example\jp{message? community?} in the test set. 
For this, we use the negative exponential of the mean cross-entropy loss 
of the language model for a given message and average over a set of messages, $M$.
\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\]
\bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
\jp{Yes}
As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models (\cref{tab:model-results}),
but this effect is different in magnitude
for models with different community embedding depths.
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two subsequent LSTM layers. 
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers.

@@

The perplexity of a CCLM gives us a measure of linguistic complexity,
but it is not a measure of how linguistically \emph{specific} the
community is. However, the CCLM induces a latent classifier,
which we call a language-model-based community classifier
(LMCC). Using the Bayes theorem, we can turn a
CCLM into a LMCC: the probability that a given message $m$
belongs to a community $c_i$ is calculated as follows:
\[P(c=c_i \mid m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m \mid c_i)$ (the probability of a message $m$
in the context of community $c_i$) is estimated by the CCLM. 
$P(m)$ is the absolute
probability of a message, which can be computed as
\[P(m) = \sum_i P(c=c_i) P(m\mid c=c_i ). \]
Since the test set is balanced across communities,
we have an even prior for $P(c=c_i)$, giving us
\[P(c=c_i\mid m) \propto  P(m\mid c_i).\]

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
%
\begin{figure*}
\begin{tikzpicture}
  \confusiongroupplot{lstm-3-1}{transformer-3-3}
\end{tikzpicture}
\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:confusion}
\end{figure*}
%
But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
noted $\Ind_j$. 
First, note that each row of $C$ is a probability distribution,
since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
whose maximum is the total number of communities (46 for us).
Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
This way $Ind_j=0$ means that the LMCC is completely certain of its
choice, while $Ind_j=1$ means that equal probability is given to all
predictions.  In general $\Ind_j$ is a score of uncertainty of the
LMCC, given messages from a community $c_j$.

@@

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinctive its language is.

\begin{table}
  \centering
  \input{floats/ind_corrs.tex}
  \caption{ Pearson's $r$ correlation coefficient between community
    indiscernibility ($\Ind_j$) and two different predictors: CCLM
    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).  }
  \label{tab:ind-corrs}
\end{table}

@@

There are, however, related applications of conditional neural language models.
\citet{Lau2017a} presents a neural language model that jointly learns to predict
words on the sentence-level and represent topics on the document level.
The topic representation is then fed back into the language model, 
improving its performance on next word prediction.
This is similar to how our model experiences improved performance
by learning community representations. 
Unlike our model, topics are inferred in an unsupervised way, 
raising the question of whether communities could be identified from 
unlabelled data as well.

Along these lines, \citet{OConnor2010} uses a Bayesian generative
model to infer communities from variation in text data.  In contrast
to our work, this model treats words as independent events, ignoring
the structure (and variation) in the construction of sequences.  It
does further suggest, however, that community-level variation could be
modeled in an unsupervised way.

@@

We cannot exclude a correlation (\cref{tab:ind-corrs}) between the
perplexity of the
% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few, and
only large communities in our test set, it is difficult to provide a
conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. With an indiscernibility $> 0.9$, the most
complex communities appear as as even mixture of the kind of language
found in all communities, to our models.

@@

The pattern of information gain by community is similar across
architectures.  Communities that benefit most from the conditioned
model behave that way for both the LSTM and transformer.  However,
there are some differences.  For example, the subreddit
\emph{relationships} has an information gain of \num{1.018} for the
best CCLM LSTM but only \num{0.998} for the best transformer model.
Why was the LSTM able to take advantage of the community information
for this community, where the transformer was not?  It would be
interesting to investigate these differences in future work, since it
could reveal differences in the kind of linguistic variation the
different model architectures capture.
